---
title: "Live Session Assignment Unit 3"
author: "Stuart Miller"
date: "5/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

I used the following commands to clone the repo:

* cd ~/projects
* git clone git@github.com:awesomedata/awesome-public-datasets.git

# Question 2

## Part a
```{r}

# path to file on local
path <- '/home/stuart/courses/Doing-Data-Science/Unit_3/awesome-public-datasets/Datasets/'
file <- 'titanic.csv/titanic.csv'

# read in csv to data.frame
df <- read.csv(paste(path,file,sep = ''))

```

## Part b

```{r}
# print a table containing the count of males and females on the titanic
print(table(df$Sex))
```


```{r}
barplot(table(df$Sex), main = 'Count of Male and Female Passengers on the Titanic',
        xlab = 'Gender', ylab = 'Count')
```

## Part c

```{r}
# calculate means for Age, Fare, and Survived variables
mean.vars <- c('Age','Fare','Survived')
apply(df[mean.vars], 2, function(x) mean(x,na.rm=T))
```

# Question 3

```{r}

# path to file on local
path <- '/home/stuart/courses/Doing-Data-Science/Unit_3/'
file <- 'sleep_data_01.csv'

# read in csv to data.frame
df <- read.csv(paste(path,file,sep = ''))

helper <- function(data){
  # calculate the following statistics by name from the data frame
  Age.median <- median(data$Age, na.rm = TRUE)
  Duration.minimum <- min(data$Duration, na.rm = TRUE)
  Duration.maximum <- max(data$Duration, na.rm = TRUE)
  RSES.sd <- sd(data$RSES, na.rm = TRUE)
  RSES.mean <- mean(data$RSES, na.rm = TRUE)
  
  # generate a report from the data
  report <- data.frame(Age.median, 
                       RSES.mean / 5, 
                       RSES.sd / 5,
                       Duration.maximum-Duration.minimum)
  
  # give descriptive names to the dataframe columns
  names(report) <- c('MedianAge','SelfEsteem','SE_SD','DurationRange')
  report
}

helper(df)

```

# Question 4

## Part a.
Used the following command to install package

* install.packages("fivethirtyeight")

## Part b.

Get the name of the 22nd dataset.

```{r}
library(fivethirtyeight)

# puts the 22nd dataset from 538 into df

a <- data(package = "fivethirtyeight")
df <-get(a$results[22,'Item'])

# print name of dataset
a$results[22,][['Item']]
```

## Part c.

```{r}
# use vignette to get PDF of package description
vignette("fivethirtyeight", package = "fivethirtyeight")

# find URL for corresponding data set name in file manually
```

The URL associated with the 22nd dataset is: http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

## Part d.

```{r}
# get dims of data in df
dim(df)
```

The output means there are 173 observations of 21 variables or 173 rows and 21 columns in this dataset.

# Question 5

## Part a.

```{r}
# get the names of the columns of df
names(df)
columns.names <- names(df)
```

```{r}
# count the number of columns in df
length(columns.names)
```

## Part b.

```{r}
#create a table from major_category
major_count <- table(df$major_category)
```


## Part c.

```{r fig1, fig.height = 10, fig.width = 10}
# create bar chart of the college majors
par(mar = c(15,4,4,2) + 0.5)
barplot(major_count, las = 2,
        main = 'Frequency of Recent College Majors',
        ylab = 'Count of Major')
mtext("College Major", side=1, line=13)
```

## Part d.

```{r}
write.csv(df,file = paste(path,'college_recent_grads.csv',sep = ''),row.names =FALSE)
```


# Question 6

Github link: https://github.com/sjmiller8182/SMU-MSDS-program














