---
title: "Simple Twitter Access with R"
output: html_notebook
---

Based on example from https://towardsdatascience.com/access-data-from-twitter-api-using-r-and-or-python-b8ac342d3efe

## Requirements

* twitteR package
* Twitter developer access token ([see link](https://towardsdatascience.com/access-data-from-twitter-api-using-r-and-or-python-b8ac342d3efe))

I have the library installed already and an access key so I will go ahead with an example. Check the link for instruction on getting an access key.

```{r}
# load the library
library(twitteR)
```

## Basic Example

**1. Authenticate with access keys.**

```{r}
# I have my access tokens stored in a csv for 'easy access'
twitterAccess <- read.csv('~/Dropbox/twitterAcess.csv')

# authenticate
with(twitterAccess, setup_twitter_oauth(as.character(APIKey),
                                        as.character(APISecretKey),
                                        as.character(AccessToken),
                                        as.character(AccessTokenSecret)))
```

**2. Query twitter search for the author of the linked example.**

* `searchTwitter` returns a data structure called 'twList'
* `twListToDF` converts a twList to a data.frame

```{r}
# query the twitter handle of author of linked example
# get 6 tweets (n = 6)
tw = searchTwitter('@GalarnykMichael', n = 5)
d = twListToDF(tw)
```

**3. Look at the available fields from the query**

```{r}
# get names of data.frame `d`
names(d)
```

**4. Get the users and associated text**

```{r}
# Note: paste auto iterates long the sequences (don't need a for loop)
print(paste(d$screenName, d$text, sep = ': '))
```
