---
title: "Live_Session_Assignment_5"
author: "Stuart Miller"
date: "6/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective

Read in data from yob2016.txt and yob2015.txt. These files contain baby names for years 2016 and 2015, along with occurence of name by gender. The data is merged along gender and name. Then the total for each name is summed and ordered by the sum. The top 10 girls names are reported to a file: top_girls_names.csv.

# Requirements

The following files are required:

* yob2016.txt
* yob2015.txt

# Question 1

## Part A

```{r}
# read in file with baby names
df <- read.csv("yob2016.txt", sep = ';', header = F)
# set column names
names(df) <- c("name","gender","count")
# convert names from factor to character
df$name <- as.character(df$name)
```

## Part B

```{r}
# look at the summary of the imported data
summary(df)
```

```{r}
# look at the structure of the imported data
str(df)
```

## Part C

```{r}
# get location of the error (obs index)
error_obs <- grep(".*yyy",df$name)
# print the error row for confirmation
print(df[error_obs,])
```

## Part D

```{r}
# remove bad observation
y2016 <- df[-error_obs,]
```

# Question 2

## Part A

```{r}
# import names for 2015
y2015 <- read.csv("yob2015.txt", header = F)
# set column names
names(y2015) <- c("name","gender","count")
# convert names from factor to character
y2015$name <- as.character(y2015$name)
```

## Part B

```{r}
tail(y2015, 10)
```

Interestingly, all the names at the end of the file start with 'Zy'. Likely this list is in alphabetical order.

## Part C

```{r}
# merge by gender then name to aviod mismatches between gender
final <- merge(y2016, y2015, by = c('gender',"name"))
names(final)[3] <- 'count_2016'
names(final)[4] <- 'count_2015'
head(final,10)
```

```{r}
# check for any NAs after merge
any(is.na(final))
```

# Question 3

## Part A

```{r}
# create new column 'Total' with sum of counts in 2015, 2016
final$Total <- final$count_2016 + final$count_2015
```

## Part B

```{r}
# sort by 'Total'
final <- final[order(-final$Total),]
head(final,10)
```

## Part C

```{r}
# only need most popular girl's names
head(final[final$gender == 'F',],10)
```

## Part D

```{r}
# get top 10 girls names without unneeded columns
top_girls_names <- head(final[final$gender == 'F',c('name','Total')],10)
# print top 10 girls names to file
write.csv(top_girls_names, file = './top_girls_names.csv', row.names = F)
```

